---
title: "Project-1"
author: "Haoran Shi and Jingyao Zhang"
date: "2024-10-17"
output: pdf_document
bibliography: references.bib
csl: apa-with-abstract.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
esg.data = read.csv("sp500_esg_data.csv")


```

# Abstract ## Tina

#Introduction ## Tina

# Data #Do Respective data

# Analysis and Results
*To see top 5% company that with lowest score, and to explore correlation between score and type of industry 
Within 90th percentile companies, pie chart of percentage each industry in the 90th percentile 
```{r plot1}
##Tina

```

*To investigate 3 industries (information technology, healthcare, financial) distributed in which percentile, category into 4, 0-25, 25-50, 50-75, 75-100
Barplot
```{r plot 2}
```

*To count categorize companies into five risk levels, negligible, low, medium, high, severe 
To see if s&p 500 companies intended to control their score under low or medium level 
Histogram: categories by risk level, frequency of companies for each risk level
See trend.
```{r plot 3}
```



# Conclusion #Jenna

# Contribution Statement

# References ## Jenna
@greenchoicenowWhatsScore
@R

